Pillar 1 - Sequential Pattern Mining

### 2019-01-28 Pillar 1 - Sequential Pattern Mining.Rmd
Runs threshold with added grades --> mkes no sense

```{r}

d_transactions <- d_course %>%

@@

  # Join with transcripts.
  inner_join(d_transcript, by = "Course ID")
```

```{r AR data prep}
# Create transactions (student level)
transactions <- d_transactions %>%
  group_by(`Student ID`) %>%
    summarize(list_course = list(`Course ID`)) %>%
  ungroup

transactions <- as(transactions$list_course, "transactions")
```

```{r AR data prep PF}
d_transactions_PF <- d_transactions %>%

  # Take outcome of course into account (pass or fail)
  mutate(PF = case_when(Grade <  5.5 ~ "fail",
                        Grade >= 5.5 ~ "pass"),
         `Course ID` = paste(`Course ID`, PF, sep = "_"))

  # Create transactions (student level)
transactions_PF <- d_transactions_PF %>%
  group_by(`Student ID`) %>%
    summarize(list_course = list(`Course ID`)) %>%
  ungroup

transactions_PF <- as(transactions_PF$list_course, "transactions")
```

## Apriori Algorithm
We apply Apriori algorithm:
```{r AR apriori}

# TODO: conmpute support RHS

AR <- apriori(
  data = transactions, 
  parameter = list(
    supp = 0, # min support
    smax = 1,    # max support
    conf = 0,    # min confidence
    minlen = 2,  # min length of rule
    maxlen = 2,  # max length of rule
    ext = TRUE,
    arem = "diff",
    aval = TRUE,
    minval = 0
    ),
  appearance = NULL,
  control = list(
    verbose = FALSE
  )
  ) %>%
  inspect

colnames(AR)[2] <- "=>"
```

Now we include data for pass and fail:
```{r apriori PF}

# creating vector of fail course
course_id_fail <- d_transactions_PF %>%
  filter(PF == "fail") %>%
  distinct(`Course ID`)

# Apply apriori
AR_PF <- apriori(
  data = transactions_PF, 
  parameter = list(
    supp = 0, # min support
    smax = 1,    # max support
    conf = 0,    # min confidence
    minlen = 2,  # min length of rule
    maxlen = 2,  # max length of rule
    ext = TRUE,
    arem = "diff",
    aval = TRUE,
    minval = 0
    ),
  appearance = list(
    both = course_id_fail$`Course ID` # only include rules with fail on LHS and RHS
    ),
  control = list(
    verbose = FALSE
    )
  ) %>%
  inspect

colnames(AR_PF)[2] <- "=>"
```

```{r AR exploration}

# TODO: MAKE APP

data_rules <- AR
data_rules <- AR_PF

# Best support
data_rules %>%
  arrange(desc(support))

# Best confidence
data_rules %>%
  filter(count >= 5) %>%
  arrange(desc(confidence))

# Best lift
data_rules %>%
  filter(count >= 5) %>%
  arrange(desc(lift))
```

@@

```{r, cache = TRUE}
tmp_distributed_transcript <- d_transcript %>%
  filter(!is.na(Grade), Grade>=0) %>%
  mutate(Round_grade = paste("g", round(Grade,0),sep=""),
         Values = T)
#This is the transcript with cumulative error gap.
distributed_transcript <- spread(tmp_distributed_transcript, key = "Round_grade", value = Values #, fill = F
               )%>%
  mutate(g1  = g0|g1,
         g2  = g1|g2,
         g3  = g2|g3,
         g4  = g3|g4,
         g5  = g4|g5,
         g6  = g5|g6,
         g7  = g6|g7,
         g8  = g7|g8,
         g9  = g8|g9,
         g10 = g9|g10,
         ) %>%
  gather(key="Gr_string", value="Gr_bool", g0, g1, g2, g3, g4,g5,g6,g7,g8,g9,g10, na.rm=T) %>%
  select(-Gr_bool )%>%
  separate(Gr_string, into=c("Trash", "Grade_new"), sep="g")%>%
  select(-Trash) %>%
  mutate(Grade= as.numeric(Grade_new))%>% #can put this into Grade and have just one grade value.
  select(-Grade_new)%>%
  arrange(`Course ID`, `Student ID`)

```
#puting everything into quasi-function

```{r}

#parameters for analysis
category <-quo(Low_grade) #Outcome or Low_grade or Discrete_grade or #None
transcript_to_use <- quo(distributed_transcript) #d_transcript or distributed_transcript 
fail_threshold <- 6.5

@@

  left_join(!!transcript_to_use, by = "Course ID") %>%
  mutate(
    Outcome = 
           case_when( Grade == -1   ~"Fail", #if we don't do this NG satisfies the condition for pass v
                      Grade  > 5.5 ~"Pass",
                               T~"Fail"),
    Low_grade =
           case_when(Grade == -1 ~"low score", #if we don't do this NG satisfies the condition for pass v
                      Grade > fail_threshold ~"high score",
                               T~"low score"),
    Discrete_grade =
           case_when(Grade <= 5.4             ~ 5.0,
                     between(Grade, 5.5, 6.4) ~ 6.0,
                     between(Grade, 6.5, 7.4) ~ 7.0,
                     between(Grade, 7.5, 8.4) ~ 8.0,
                     between(Grade, 8.5, 9.4) ~ 9.0,
                     Grade >= 9.5             ~ 10.0,
                     T ~ -99999    #this is a control, we should not have any grades here
                     ),
    None= "" # this might not be optimal
         ) %>%

@@

>>>>

```{r AR data prep}
d_transactions <- d_course %>%

@@

  # Join with transcripts.
  left_join(d_transcript, by = "Course ID") %>%

  # Create transactions (student level)
  group_by(`Student ID`) %>%
    summarize(list_course = list(`Course ID`)) %>%
  ungroup

transactions <- as(d_transactions$list_course, "transactions")
```

## Apriori Algorithm
We apply Apriori algorithm:
```{r}
results <- apriori(
  data = transactions, 
  parameter = list(
    supp = 0.02,
    smax = 1, # max support
    conf = 0.4,
    minlen = 2,
    maxlen = 2
    ),
  appearance = NULL,
  control = NULL
  ) %>%
  inspect

colnames(results)[2] <- "association"

results %>%
  arrange(desc(confidence))
```
Now we include data for pass and fail:
```{r}
d_transaction_passed <- d_course %>%
  select(Type,`Course ID`, Letters) %>%
  filter(Type == "Elective",
         ! Letters %in% c("SKI", "PRO"),
         ! Letters %in% c("SAS", "SAH", "SAC")) %>%
  select(`Course ID`) %>%
  left_join(d_transcript, by = "Course ID") %>%
  mutate(Outcome = case_when(Grade=="NG"~"Fail",
                             Grade > 5.5 ~"Pass",
                             T~"Fail")) %>%
  mutate(Item= paste(`Course ID`, Outcome, sep = "_")) %>%
  select(`Student ID`, Item) %>%
  group_by(`Student ID`) %>%
  summarize(list_course = list(Item)) %>%
  group_by()

transactions_passed <- as(d_transaction_passed$list_course, "transactions")

```
We apply Apriori algorithm:
```{r}
results_passed <- apriori(
  data = transactions_passed, 
  parameter = list(
    supp = 0,
    smax = 1, # max support
    conf = 0.4,
    minlen = 2,
    maxlen = 2
    ),
  appearance = NULL,
  control = NULL
  ) %>%
  inspect

colnames(results_passed)[2] <- "association"

results_passed %>%
  separate(lhs, into = c("LHS", "LHS_pass"), sep= "_", remove = F) %>%
  separate(rhs, into = c("RHS", "RHS_pass"), sep= "_", remove = F) %>%
  filter(RHS_pass == "Fail}") %>%
  arrange(desc(count))
```

@@

```{r, cache = TRUE}
tmp_distributed_transcript <- d_transcript %>%
  filter(!is.na(Grade), Grade>=0) %>%
  mutate(Round_grade = paste("g", round(Grade,0),sep=""),
         Values = T)
#This is the transcript with cumulative error gap.
distributed_transcript <- spread(tmp_distributed_transcript, key = "Round_grade", value = Values #, fill = F
               )%>%
  mutate(g1  = g0|g1,
         g2  = g1|g2,
         g3  = g2|g3,
         g4  = g3|g4,
         g5  = g4|g5,
         g6  = g5|g6,
         g7  = g6|g7,
         g8  = g7|g8,
         g9  = g8|g9,
         g10 = g9|g10,
         ) %>%
  gather(key="Gr_string", value="Gr_bool", g0, g1, g2, g3, g4,g5,g6,g7,g8,g9,g10, na.rm=T) %>%
  select(-Gr_bool )%>%
  separate(Gr_string, into=c("Trash", "Grade_new"), sep="g")%>%
  select(-Trash) %>%
  mutate(Grade= as.numeric(Grade_new))%>% #can put this into Grade and have just one grade value.
  select(-Grade_new)%>%
  mutate(Item=paste(`Course ID`,Grade, sep="_"))%>%
  select(`Student ID`,`Course ID`)%>%
  arrange(`Course ID`, `Student ID`) %>%

```
#puting everything into quasi-function
--------------------------------------------------
```{r}

#parameters for analysis
category <-quo(Low_grade) #Outcome or Low_grade or Discrete_grade or #None
transcript_to_use <- "distributed_transcript" #d_transcript or distributed_transcript 
fail_threshold <- 6.5

@@

  left_join(get(transcript_to_use), by = "Course ID") %>%
  mutate(
    Outcome = 
           case_when( Grade == -1   ~"Fail", #if we don't do this NG satisfies the condition for pass v
                      Grade  > 5.5 ~"Pass",
                               T~"Fail"),
    Low_grade =
           case_when(Grade == -1 ~"low score", #if we don't do this NG satisfies the condition for pass v
                      Grade > fail_threshold ~"high score",
                               T~"low score"),
    Discrete_grade =
           case_when(Grade <= 5.4             ~ 5.0,
                     between(Grade, 5.5, 6.4) ~ 6.0,
                     between(Grade, 6.5, 7.4) ~ 7.0,
                     between(Grade, 7.5, 8.4) ~ 8.0,
                     between(Grade, 8.5, 9.4) ~ 9.0,
                     Grade >= 9.5             ~ 10.0,
                     T ~ -99999    #this is a control, we should not have any grades here
                     ),
    None= "" # this might not be optimal
         ) %>%

@@

#checking results
as(transactions, "data.frame") #WARNING: THIS TAKES A LOT OF TIME, FOR SOME REASON
```
##Create filter condition.
```{r}
string_filter <- case_when(quo_name(category)=="Outcome"        ~ "Fail",
                           quo_name(category)=="Low_grade"      ~ "low score",
                           quo_name(category)=="Discrete_grade" ~ "5",
                           quo_name(category)=="None"           ~ "_" #useless filter
                           )
```
##CSPADE Algorithm:
```{r}
frequent_trajectories <- cspade(transactions, parameter = list(support = 0,
                                                               maxsize = 1,
                                                               maxlen = 2 #,
                                                               #mingap=, #gap between consecutive elements
                                                               #maxgap=, #hap between consecutive elements
                                                               #maxwin=  #gap between ANY two elements
                                                                 ), control   = list(verbose = TRUE))

  filter(str_detect(RHS, string_filter)) %>% #FILTER (YES) --> uncomment, (NO)-->comment

```
here I'l playing around with the correct confidence:
#correcting lookup for this case:
First we find out raw proportions per course (how many were low grade/high grade, out of a particular course, and the percentage this represents from the total)
```{r}
lookup_table <- d_course %>%

  select(`Student ID`, #Item,  #************************************ <- 
         `Course ID`, !!category, Time) %>% # Course ID and Time may not be necessary

```{r}
n_students <- length(unique(d_sequence_transactions$`Student ID`))

sequence_rules <- pass_rules_df %>%
  separate(LHS, into=c("LHS Course ID", "LHS Score"), sep="_", remove= FALSE) %>%
  separate(RHS, into=c("RHS Course ID", "RHS Score"), sep="_", remove= FALSE) %>%
  mutate(`RHS Course ID`= substr(`RHS Course ID`, start=4, stop=10))%>% #add more
  left_join(lookup_table, by = c("RHS Course ID"="Course ID")) %>%

  # Compute statistics
  group_by(LHS,`RHS Course ID`) %>%
  arrange(LHS,`RHS Course ID`)
  mutate(
    Prob_both = sum(support), 
    Conf_corr = support/Prob_both,
    n_rule    = support * n_students,
    n_both    = Prob_both * n_students,
    lift_diff = Conf_corr - Prop_low,
    lift_corr = Conf_corr / Prop_low
    ) %>%
  ungroup() %>%

---------------------------------------------------
